### 2024-08-12 rewrite.qmd

---
title: "Conceptualizing open health data platforms for low- and middle income countries"
format:
  elsevier-html:
    toc: true
    toc-depth: 3
    fig-cap-location: top
    code-fold: true
  elsevier-pdf:
    keep-tex: true
    journal:
      name: JIMR
      cite-style: authoryear
author:
  - name: Daniel Kapitan
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands
      - name: Eindhoven University of Technology
        city: Eindhoven
        country: the Netherlands
    orcid: 0000-0001-8979-9194
    email: daniel@kapitan.net
  - name: Femke Heddema
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands
  - name: Julie Fleischer
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands
  - name: Chris Ihure
    affiliations:
      - name: PharmAccess Kenya
        city: Nairobi
        country: Kenya
  - name: Steven Wanyee
    affiliations:
      - name: IntelliSOFT
        city: Nairobi
        country: Kenya
  - name: Alessandro Pietrobon 
    affiliations:
      - name: ONA
        city: Nairobi
        country: Kenya
  - name: Ryan Chrichton
    affiliations:
      - name: Jembi Health Systems
        city: Durban
        country: South Africa
  - name: John Grimes
    affiliations:
      - name: Australian e-Health Research Centre
        city: Brisbane
        country: Australia
  - name: Paula van Brakel
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands

  - name: Mark van der Graaf
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands
  - name: Nicole Spieker
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands

abstract: |
  TO DO: add abstract.
keywords: [Analytics-on-FHIR, SQL-on-FHIR, HIE, data platforms, LMICs, digital health]
reference-section-title: References
bibliography: pharmaccess.bib
---

## Introduction {#sec-introduction}

Digital transformation have the potential to make inclusive health systems in low- and middle income countries (LMICs) a reality by connecting patients and professionals, using data as a currency of trust. Adoption of digital technologies in healthcare means more than simply increasing efficiency or cutting costs. Instead, digital technologies have the potential to i) increase the availability, accessibility, acceptability, and quality of health services as we know them; ii) change nature and direction of health care, and the possibility of making it more preventive, personalised, and mobile through the use of such technologies; and iii) enfranchise patients and communities (and particularly vulnerable groups including children and young people) and evolve their relationship with health professionals and providers, thus helping shape the health system according to the needs of the patients and communities [@kickbusch2021lancet].

Yet, to date the global digital health ecosystem is still project-centric, resulting in data fragmentation and technology lock-in, compromising health care delivery [@mehl2023fullstac]. There is no dearth of the number of digital health interventions: sub-Saharan Africa has seen over 700 such projects in the past 10 years. There is, however, a fundamental lack in coordination, integration, scalability, sustainability, and equitable distribution of investments in digital health. Health policymakers and the global health community at large need to urgently institute coordination mechanisms to terminate unending duplication and disjointed vertical implementations and manage solutions for scale [@karamagi2022ehealth].

In this context it is essential to have a good understanding of digital platforms and their associated ecosystem as an essential enabling socio-technical infrastructure upon  which digital health interventions can be realized in a more standardized, scaleable and sustainable fashion. Digital platforms are a challenging research object because of their distributed nature and intertwinement with institutions, markets and technologies [@dereuver2018digital]. Digital platform ecosystems vary according to three core building blocks: (1) platform ownership, (2) value-creating mechanisms, and (3) complementor autonomy [@hein2020digital]. Digital platforms have disrupted many sectors but have just started to make inroads into highly regulated industries such as healthcare [@ozalp2022digital].

In this light, the challenges faced by LMICs in establishing national digital health platforms have a lot in common with those faced by high income countries. From a technological perspective, interoperability issues, weak integrations, siloed data repositories and overall lack of openness are often reported as key impediments [@malm-nicolais2023exploring;@mehl2023fullstac]. From a societal perspective, issues pertaining to the winner-takes-all nature of digital platforms are hotly debated as many jurisdictions make work to ensure these new digital health platforms indeed serve the common good of achieving universal health coverage [@sharon2018when].