### 2024-08-13 index.tex
wip

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{JIMR}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Conceptualizing open health data platforms for low- and middle income countries},
  pdfauthor={Daniel Kapitan; Femke Heddema; Julie Fleischer; Chris Ihure; Steven Wanyee; Alessandro Pietrobon; Ryan Chrichton; John Grimes; Paula van Brakel; Mark van der Graaf; Nicole Spieker},
  pdfkeywords={Analytics-on-FHIR, SQL-on-FHIR, HIE, data
platforms, LMICs, digital health},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Conceptualizing open health data platforms for low- and middle
income countries}
\author[1,2]{Daniel Kapitan%
%
}
 \ead{daniel@kapitan.net} 
\author[1]{Femke Heddema%
%
}

\author[1]{Julie Fleischer%
%
}

\author[3]{Chris Ihure%
%
}

\author[4]{Steven Wanyee%
%
}

\author[5]{Alessandro Pietrobon%
%
}

\author[6]{Ryan Chrichton%
%
}

\author[7]{John Grimes%
%
}

\author[1]{Paula van Brakel%
%
}

\author[1]{Mark van der Graaf%
%
}

\author[1]{Nicole Spieker%
%
}

\affiliation[1]{organization={PharmAccess
Foundation},city={Amsterdam},country={the
Netherlands},countrysep={,},postcodesep={}}
\affiliation[2]{organization={Eindhoven University of
Technology},city={Eindhoven},country={the
Netherlands},countrysep={,},postcodesep={}}
\affiliation[3]{organization={PharmAccess
Kenya},city={Nairobi},country={Kenya},countrysep={,},postcodesep={}}
\affiliation[4]{organization={IntelliSOFT},city={Nairobi},country={Kenya},countrysep={,},postcodesep={}}
\affiliation[5]{organization={ONA},city={Nairobi},country={Kenya},countrysep={,},postcodesep={}}
\affiliation[6]{organization={Jembi Health
Systems},city={Durban},country={South
Africa},countrysep={,},postcodesep={}}
\affiliation[7]{organization={Australian e-Health Research
Centre},city={Brisbane},country={Australia},countrysep={,},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
TO DO: add abstract.
\end{abstract}

\begin{keyword}
    Analytics-on-FHIR \sep SQL-on-FHIR \sep HIE \sep data
platforms \sep LMICs \sep 
    digital health
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{sec-intro}

\subsection{The paradox of open for digital vs.~data platforms in
healthcare}\label{the-paradox-of-open-for-digital-vs.-data-platforms-in-healthcare}

It is a widely held belief that digital technologies have an important
role to play in strengthening health systems in low- and middle income
countries (LMICs), as exemplified by the WHO global strategy on digital
health \citep{who2021global}. The adoption rate of mobile phones in
LMICs has been an important driver in implementing digital health
solutions \citep{mccool2022mobile}. Yet, there are many shortcomings and
challenges, including the current fragmentation of digital platforms and
the lack of clear-cut pathways of scaling up digital health programmes,
such that they can support sustainable and equitable change of national
health systems in LMICs
\citep{mehl2023fullstac, mccool2022mobile, who2019recommendations, neumark2021digital}.

A commonly used perspective to scrutinize digital health is to consider
it as a digital platform \citep{dereuver2018digital}. Digital platforms
have disrupted many sectors but have just started to make inroads into
highly regulated industries such as healthcare \citep{ozalp2022digital}.
In this light, the challenges faced by LMICs in establishing national
digital health platforms have a lot in common with those faced by high
income countries. From a technological perspective, interoperability
issues, weak integrations, siloed data repositories and overall lack of
openness are often reported as key impediments
\citep{malm-nicolais2023exploring, mehl2023fullstac}. From a societal
perspective, issues pertaining to the winner-takes-all nature of digital
platforms are hotly debated as many jurisdictions make work to ensure
these new digital health platforms indeed serve the common good of
achieving universal health coverage \citep{sharon2018when}.

Case studies on digital platforms in healthcare point to an emerging
pattern where the focus shifts from the digital platform with its
defining software and hardware components, to the data as the primary
object of interest in and of itself
\citep{ozalp2022digital, alaimo2022organizations}. This observation ties
into with the proposed research agenda by de Reuver et al.~to consider
data platforms as a phenomenon distinct from digital platforms
\citep{dereuver2022openness}. Generally, data platforms inherit the
characteristics of digital platforms. From an economic perspective, for
example, both types exhibit multi-sided markets. At the same time, data
platforms differ as their main offerings revolve around data. From an
ecosystem perspective, data platforms have more moderate network effects
and are more susceptible to fragmentation and heterogeneity
\citep{dereuver2022openness}.

Particularly relevant in the context of open health data platforms
(OHDPs), is the conceptualization of openness. The shift in perspective
from digital platforms to data platforms coincides with the paradox of
open \citep{keller2021paradox}. Originally, openness of digital
platforms focused on open source, open standards and copyrights, which
by has been superseded by ``\ldots{} conflicts about privacy, economic
value extraction, the emergence of artificial intelligence, and the
destabilizing effects of dominant platforms on (democratic) societies.
Instead of access to information, the control of personal data has
emerged in the age of platforms as the critical contention.''
\citep{keller2021paradox}. These conflicts are particularly salient in
the healthcare domain, where people are generally willing to share their
health data to receive the best care (primary use), while the attitude
towards secondary use of health data varies greatly depending on the
type and context \citep{cascini2024health}. The shift in perspective
from digital platforms supporting primary data sharing toward data
platforms supporting secondary data sharing is one of the key issues
surrounding the polemic of data spaces \citep{otto2022designing} and
data solidarity
\citep{kickbusch2021lancet, prainsack2022data, prainsack2023beyond, purtova2023data}.
Openness is particularly relevant if we are to realize a
solidarity-based approach to health data sharing that i) gives people a
greater control over their data as active decision makers; ii) ensures
that the value of data is harnessed for public good; and iii) moves
society towards equity and justice by counteracting dynamics of data
extraction \citep{prainsack2022data}.

\subsection{From health information exchanges to health data
platforms}\label{from-health-information-exchanges-to-health-data-platforms}

This paper is motivated by the conflation of a number of developments
relevant to the design and implementation of open, solidarity-based
OHDPs in LMICs. First, the OpenHIE framework \citep{openhie} has been
adopted by many sub-Saharan African countries \citep{mamuye2022health}
as the architectural blueprint for implementing nation-wide health
information exchanges (HIE), including Nigeria \citep{dalhatu2023paper},
Kenya \citep{thaiya2021adoption} and Tanzania \citep{nsaghurwe2021one}.
These countries have, as a matter of course, extended the framework to
include ``data \& analytics services'' as an additional domain. The
rationale for this addition is to facilitate secondary reuse of health
data for academic research, real-world evidence studies etc. which can
be framed within the context of ongoing efforts towards Findable,
Accessible, Interoperable and Reusable (FAIR) sharing of health data
\citep{guillot2023fair}. In doing so, however, we have implicitly moved
from conceptualizing digital health platforms for primary data sharing
(the original OpenHIE specification) to health data platforms for
secondary data sharing. This is problematic because the notion of
openness, which is assumed to be essential in establishing
solidarity-based approaches to data sharing, is inherently different for
a data platform compared to a digital platform.

Conceptually, the OpenHIE framework constitutes a framework for an open
digital platform. Openness for digital platforms refers to i) the use of
open boundary resources, that is, specifications for the various
healthcare specific workflows and information standards such as FHIR;
and ii) the use of open source components that are available as digital
public goods \citep{digitalpublicgoods}. If we are to use the OpenHIE
framework as an open data platform, we need to extend the standards,
technologies and architecture to include functionality for secondary
data sharing and reuse. The lack of detailed specifications and
consensus of this addition to OpenHIE currently stands in the way of
development projects that aim to establish OHDPs in LMICs. The purpose
of this paper, therefore, is to specificy how new standards and
technologies can be integrated into the OpenHIE architecture such that
an open OHDP can be realized that supports the following different types
of data sharing.

\subsubsection{1. Sharing of data sets at the most granular (patient)
level, persisted as longitudinal
records}\label{sharing-of-data-sets-at-the-most-granular-patient-level-persisted-as-longitudinal-records}

The Shared Health Record (SHR) in OpenHIE provides an operational,
real-time transactional data source which is intended for primary data
sharing. The specification explicitly states that the SHR is distinct
from a datawarehouse. At the same time, the Health Management
Information System (HMIS) component is specified to fulfil functional
requirements that are typically provided by datawarehouses.

\begin{itemize}
\tightlist
\item
  \textbf{Q1:} How can proven solution designs and modern technologies
  from data warehousing and engineering be integrated in the OpenHIE
  specification and how does this relate to the curren specifications of
  the SHR and HMIS.
\end{itemize}

\subsubsection{2. Sharing of data products derived from the original
data
set}\label{sharing-of-data-products-derived-from-the-original-data-set}

OpenHIE specifies that the Health Management Information System (HMIS)
component should support a workflow to validate and save aggregate data
based on the emerging
\href{https://www.ihe.net/uploadedFiles/Documents/QRPH/IHE_QRPH_Suppl_ADX.pdf}{IHE
Aggregate Data Exchange (ADX) standard}. This is just one specific
workflow within a larger data analytics value chain that covers i)
collection, ii) standardization, iii) cleaning, iv) storage, v) analysis
and vi) distribution \citep{curry2016big}. Sharing of benchmarking, and
decision-support tools based on various statistical (machine learning)
models that have been trained on data sets and/or data product, can be
shared for standalone use.

\begin{itemize}
\tightlist
\item
  \textbf{Q2:} How can proven solutions designs from the data
  warehousing and engineering community be included in OpenHIE to
  support sharing of wider variety of data products.
\end{itemize}

\subsubsection{3. Sharing of secure computational environments to access
and work with the
data}\label{sharing-of-secure-computational-environments-to-access-and-work-with-the-data}

Current solution design of centralized data storage is increasingly
being challenged. Recent work points to federated learning (FL)
\citep{rieke2020future} and privacy- enhancing technologies (PETs)
\citep{scheibner2021revolutionizing, jordan2022selecting} as a better
long-term solution to secure and equitable sharing of data. Given the
resource constrained environments of LMICs, is is particularly
challenging to devise a solution that is economically feasible and
\ldots{} {[}TO DO: add feedback from Mark vdG{]}

Federated solutions:

\begin{itemize}
\tightlist
\item
  KETOS OMOP-FHIR \citep{gruendner2019ketos}
\item
  Personal Health Train on FHIR \citep{choudhury2020personal}
\item
  OHDSI analytics \citep{khalid2021standardized}
\item
  CODA project \citep{mullie2023coda}
\item
  GenoMed4All, in evalution phase for -omis \citep{cremonesi2023need}
\end{itemize}

Similar approaches taken in the context of data sharing for pandemic
response, for example:

\begin{itemize}
\item
  VODAN Africa, based on electronic data capture with templates
  \citep{purnamajati2022data}
\item
  Global Data Sharing Initiative (GDSI) \citep{pirmani2023journey}
\item
  \textbf{Q3:} How can a solution be designed that is economically
  viable, can be deployed in a bottom-up fashion, support data
  governance principles \ldots{}
\end{itemize}

\section{Methods}\label{methods}

In this paper, we present a design that extends the OpenHIE
specification to include the three types of data sharing mentioned
above. Using the full-STAC approach \citep{mehl2023fullstac} we combine
open standards, open technologies and open architectures into a coherent
modular OHDP that can be configured and reused across a variety of
use-cases. We employ a formative, naturalistic evaluation to assess the
technical risk and efficacy of the design \citep{venable2016feds}. Given
that it is prohibitively expensive to evaluate the design a real-world
setting, we aim to minimize technological risks and maximize the
efficacy of the design by considering three examples of health data
platforms in LMICs, namely:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  the OpenHIM platform
  (\url{https://jembi.gitbook.io/openhim-platform/})
\item
  the OnaData platform
  \url{https://ona.io/home/products/ona-data/features/}
\item
  the work conducted at PharmAccess Foundation as part of the MomCare
  programme (\url{https://health-data-commons.pharmaccess.org}).
\end{enumerate}

These three real-world implementations are evaluated along the following
dimensions:

\begin{itemize}
\tightlist
\item
  What is the level of openness of the implementation, specifically in
  terms of the three types of data sharing and platform-to-platform
  opennenss?
\item
  How are the core maintenance functions of the OHDP implemented, where
  we particularly focus on functions related to metadata management,
  data lineage and access control mechanisms?
\item
  Is the solution suitable for downward scalability, where we focus on
  the computer and storage requirements of the solution and assess the
  minimal requirements for on-site resources (computers, edge devices)
\end{itemize}

As part of our design research, we have taken a narrative approach in
surveying existing scientific studies on health data platforms, focusing
on the seminal reports and subsequently searching forward citations. In
addition, we have searched the open source repositories (most notably
GitHub) and the online communities (OpenHIE community, FHIR community)
to search for relevant open standards, technologies and architectures.
This paper should not be considered as a proper systematic review.

The main contributions of this paper are i) description of a framework
for the components of the Data \& Analysis Services that builds on
current best practices from the data engineering community into the
OpenHIE framework; ii) evaluation of different implementations and
design options for type 1 and 2 data sharing within an extended OpenHIE
architecture; and iii) provide open content for the MomCare
implementation to facilitate adaptation and deployment. As such, it aims
to inform future developments and implementation of open digital health
platforms in LMICs.

\section{Design}\label{design}

\subsection{Open standards: using FHIR as the common data
model}\label{open-standards-using-fhir-as-the-common-data-model}

The recent convergence to FHIR as the de facto standard for information
exchange has fuelled the development of OpenHIE. FHIR is currently used
both for routine healthcare settings
\citep{amar2024electronic, ayaz2021fast} and clinical research settings
\citep{duda2022hl7, vorisek2022fast} and is increasingly being used in
LMICs as well. The guidelines and standards of the African Union
explicitly state FHIR is to be used as the messaging standard
\citep{2023african}. The FHIR-native OpenSRP platform
\citep{mehl2020open} has been deployed in 14 countries targeting various
patient populations, amongst which a reference implementation of the WHO
antenatal and neonatal care guidelines for midwives in Lombok, Indonesia
\citep{summitinstitutefordevelopment2023bunda, kurniawan2019midwife}. In
India, FHIR is used as the underlying technology for the open Health
Claims Exchange protocol specification, which has been adopted by the
Indian government as the standard for e-claims handling \citep{hcx}.
This range of utilizations showcase the standards' widespread
applicability. The proceedings of the OpenHIE conference 2023 attest to
the fact that FHIR and open source technologies are embraced as critical
enablers in implementing health information exchanges in LMICs
\citep{ohie2023unconference}.

Various studies have investigated the merits of FHIR and its performance
vis-a-vis other healthcare standards. Comparisons between OpenEHR, ISO
13606, OMOP and FHIR have been made
\citep{ayaz2023transforming, mullie2023coda, rinaldi2021openehr, cremonesi2023need, sinaci2023data, tsafnat2024converge}.
A study involving 10 experts comparing OpenEHR, ISO 13606 and FHIR
concluded that i) these three standards are functionally and technically
compatible, and therefore can be used side by side; and that ii) each of
these standards have their strengths and limitations that correlate with
their intended use as summarized in the Table~\ref{tbl-comparison}.

\begin{table}

\caption{\label{tbl-comparison}Comparison of OpenEHR, ISO 13606 and FHIR
standards}

\centering{

\includegraphics{images/comparison-ehr-standards.png}

}

\end{table}%

For an infectious diseases dataset with a limited scope, OpenEHR, OMOP
and FHIR have been compared and found all to be equally suitable
\citep{rinaldi2021openehr}. Comparing OMOP and FHIR, the latter has been
found to support more granular mappings required for analytics and was
therefore chosen as the standard for the CODA project
\citep{mullie2023coda}.

Although FHIR was originally designed only for exchange between systems,
we propose to use it as the common data model for the design presented
here for the following reasons:

\begin{itemize}
\tightlist
\item
  Industry adoption has significantly increased, as exemplified by
  FHIR-based offering by major cloud providers such as Google, Azure and
  AWS. Also, Africa CDC has explicitly chosen FHIR as the preferred
  standard;
\item
  The widespread availibility of the Bulk FHIR API
  \citep{mandl2020push, jones2021landscape} enables bulk, file-based
  batchwise processing for analytics using the lakehouse architecture as
  detailed in the next section;
\item
  The concept of FHIR Profiles allow localisation to tailor the standard
  to a specific use case. A profile defines rules, extensions, and
  constraints for a resource. We posit that the possible penalty of this
  flexibility, namely having to manage different FHIR versions and/or
  profiles, is less of an issue in the context of LMICs where first
  priority is to exchange datasets such as the International Patient
  Summary (IPS) that are less complex compared to the requirements for
  high income countries;
\item
  Being based on webstandards, the FHIR standard lends itself best for
  further separation of concerns as envisioned by the composable data
  stack. This is an important enabler for the downward scalability of
  the solution;
\item
  With its inherent, graph-like nature, FHIR can be readily incorporated
  into the principles of FAIR data sharing, where FHIR-based data
  repositories can be integrated in an overarching netwerk of FAIR data
  stations \citep{sinaci2023data, pedrera-jimenez2023can}.
\end{itemize}

TO DO: re-use these elements from table 1 into this section:

\begin{itemize}
\tightlist
\item
  Given that the Bulk FHIR API has by now been incorporated in all major
  FHIR implementations \citep{mandl2020push, jones2021landscape}, the
  FHIR standard can be readily used to bridge the gap between primary
  (transactional) and secondary (analytics) data sharing.
\item
  \textbf{SQL-on-FHIR specification} \citep{sql-on-fhir}: provides a
  standardised approach to make FHIR work well with familiar and
  efficient SQL engines that are most commonly used in analytical
  workflows. Builds on FHIRPath \citep{hl72020fhirpath} expressions in a
  logical structure to specify things like column names and unnested
  items. Implementations of this approach are available or forthcoming,
  including open source implementations such as Pathling
  \citep{grimes2022pathling} and commercial offerings like Aidbox
  \citep{aidbox}. Modern data platforms take this Within the HL7
  ecosystem much progress has been made to support ``analytics-on-FHIR''
  with standards including \href{https://hl7.org/fhirpath/}{FHIRPath}
  and the new SQL-on-FHIR specification \citep{sql-on-fhir}. Given the
  use of FHIR as the common data model (the rationale of which will be
  described later), we formulate the second design question:
\end{itemize}

Possible risks pertaining to the use of FHIR as the common data model,
most notably the possible incompatibilities and/or high costs of
maintenance in supporting different versions, will be addressed in the
Discussion.

\subsection{Open architecture: extending OpenHIE with a composable data
stack}\label{open-architecture-extending-openhie-with-a-composable-data-stack}

\subsubsection{Evolving the data lakehouse to a composable data
stack}\label{evolving-the-data-lakehouse-to-a-composable-data-stack}

Data management and analytics platforms have undergone significant
changes since the first generation of data warehouses were introduced.
Recent studies have shown that the current practice has converged
towards the lakehouse as one of the most commonly used solution designs
\citep{armbrust2021lakehouse, hai2023data, harby2022data}. Lakehouses
typically have a zonal architecture that follow the
Extract-Load-Transform pattern (ELT) where data is ingested from the
source systems in bulk (E), delivered to storage with aligned schemas
(L) and transformed into a format ready for analysis (T)
\citep{hai2023data}. The discerning characteristic of the lakehouse
architecture is its foundation on low-cost and directly-accessible
storage that also provides traditional database management and
performance features such as ACID transactions, data versioning,
auditing, indexing, caching, and query optimization
\citep{armbrust2021lakehouse}. Lakehouses thus combine the key benefits
of data lakes and data warehouses: low-cost storage in an open format
accessible by a variety of systems from the former, and powerful
management and optimization features from the latter.

With respect to current implementations of lakehouse data platforms, we
observe a proliferation of tools with as yet limited standards to
improve technical interoperability. In the analysis of Pedreira et al.
\citep{pedreira2023composable} the requirement for specialization in
data management systems has evolved faster than our software development
practices. This situation has created a siloed landscape composed of
hundreds of products developed and maintained as monoliths, with limited
reuse between systems. It has also affected the end users, who are often
required to learn the idiosyncrasies of dozens of incompatible SQL and
non-SQL API dialects, and settle for systems with incomplete
functionality and inconsistent semantics. To remedy this, Pedreira et
al.~call to (re-)design and implement modern data platforms in terms of
a `composable data stack' as a means to decrease development and
maintenance cost and pick-up the speed of innovation.

While the lakehouse architecture separates the concerns of compute and
storage, the composable data stack takes the separation of concerns is
taken one step further. A composable data system
(Figure~\ref{fig-composable-data-stack}), not only separates the storage
(layer 3) and execution (layer 2), but also separates the user interface
(layer 1) from the execution engine by introducing standards including
Substrait for Intermediate Representation (standard A, IR) and Apache
Arrow for data connectivity and data memory layout (standards B and C,
respectively). The first generation of open source components are
already available. For example, the Ibis user interface is currently
sufficiently mature to offer a standardized dataframe interface to 19
different execution engines \citep{ibisproject}.

\begin{figure}

\begin{minipage}{\linewidth}

\begin{figure}[H]

{\centering \includegraphics{images/composable-data-stack.png}

}

\subcaption{a)}

\end{figure}%

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\begin{figure}[H]

{\centering \includegraphics{images/composable-data-stack-implementation-2.png}

}

\subcaption{b)}

\end{figure}%

\end{minipage}%

\caption{\label{fig-composable-data-stack}Schematic overview of the
composable data stack showing a) the overall architecture and b)
examples of implementations. Images taken from
\url{https://voltrondata.com/codex}.}

\end{figure}%

\subsubsection{SQL-on-FHIR v2 as an intermediate representation for FHIR
data in tabular
format}\label{sql-on-fhir-v2-as-an-intermediate-representation-for-fhir-data-in-tabular-format}

The premise of separating the user interface from the execution engine
is directly related to the key objective of the SQL-on-FHIR project
(\url{https://build.fhir.org/ig/FHIR/sql-on-fhir-v2/}), namely to make
large-scale analysis of FHIR data accessible to a larger audience,
portable between systems and to make FHIR data work well with the best
available analytic tools, regardless of the technology stack. However,
to use FHIR effectively analysts require a thorough understanding of the
specification as FHIR is represented as a graph of resources, with
detailed semantics defined for references between resources, data types,
terminology, extensions, and many other aspects of the specification.
Most analytic and machine learning use cases require the preparation of
FHIR data using transformations and tabular projections from its
original form. The task of authoring these transformations and
projections is not trivial and there is currently no standard mechanisms
to support reuse.

The solution of the SQL-on-FHIR project is to provide a specification
for defining tabular, use case-specific views of FHIR data. The view
definition and the execution of the view are separated, in such a way
that the definition is portable across systems while the execution
engine (called runners) are system-specific tools or libraries that
apply view definitions to the underlying data layer, optionally making
use of annotations to optimize performance.

\subsubsection{Extending OpenHIE with a FHIR-based composable data
stack}\label{extending-openhie-with-a-fhir-based-composable-data-stack}

We propose to extend the OpenHIE architecture with a ``Data and
Analytics Services'' domain with different service layers by
synthesizing the current best practices of the a lakehouse architecture
of \citep{hai2023data, harby2022data, harby2024data} and the composable
data stack \citep{pedreira2023composable} (Figure~\ref{fig-ohie},
Table~\ref{tbl-data-and-analysis-services}). These 5 services are
considered the core of the FHIR data platform, and although in practice
often a downstream dashboarding or visualization component is used, that
component is not the main focus of our analysis. Rather, we aim to
elucidate and conceptualize the core ``FHIR Data Lakehouse''.

\begin{figure}

\centering{

\includegraphics{./images/openhie-extended-architecture.png}

}

\caption{\label{fig-ohie}Proposed extension of the OpenHIE architecture
that includes ``Data and Analytics Services'' as an additional service
domain.}

\end{figure}%

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1852}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8148}}@{}}
\caption{Definition of Data and Analysis
Services}\label{tbl-data-and-analysis-services}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Service
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functional requirements
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Service
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functional requirements
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Ingestion & \begin{minipage}[t]{\linewidth}\raggedright
\begin{itemize}
\tightlist
\item
  Bulk
\item
  Streaming
\end{itemize}
\end{minipage} \\
Storage & \begin{minipage}[t]{\linewidth}\raggedright
\begin{itemize}
\tightlist
\item
  File-based blob storage
\item
  Database optimized for online analytical processing (OLAP)
\end{itemize}
\end{minipage} \\
Maintenance & \begin{minipage}[t]{\linewidth}\raggedright
\begin{itemize}
\tightlist
\item
  SQL-on-FHIR View defintions
\item
  Catalog and other maintenance-related functions as defined by Hai et
  al.
\end{itemize}
\end{minipage} \\
Processing \& API & \begin{minipage}[t]{\linewidth}\raggedright
\begin{itemize}
\tightlist
\item
  SQL-on-FHIR Runner
\item
  Execution engine on tabular data as defined in composable data stack
\item
  Capability to participate as a node in federated learning / MPC
  network
\item
  Read-only access to storage
\end{itemize}
\end{minipage} \\
Data Consumer & \begin{minipage}[t]{\linewidth}\raggedright
\begin{itemize}
\tightlist
\item
  SQL interactive development environment (IDE)
\item
  Interactive notebook computing environment \citep{granger2021jupyter}
\item
  BI reporting, dashboarding and data visualization
\end{itemize}
\end{minipage} \\
\end{longtable}

\paragraph{Ingestion}\label{ingestion}

Default workflow is extraction of data from SHR using Bulk FHIR API.
Data contains metadata (incl.~FHIR versions) and fully qualified
semantics, for example, coding systems. Despite this, metadata
extraction and metadata modeling is still required to meet the FAIR
requirements. Issues that need to be solved by these services:

\begin{itemize}
\tightlist
\item
  To prepare for future updates of FHIR versions
\item
  Implement late-binding principle of having increasingly more specific
  FHIR profiles as bulk FHIR data propagates through lakehouse
\end{itemize}

\paragraph{Storage}\label{storage}

\begin{itemize}
\tightlist
\item
  File-based:

  \begin{itemize}
  \tightlist
  \item
    from ndjson to parquet
  \item
    possibly used delta lake for time versioning
  \item
    separation of storage from compute not only for benefits of lower
    TCO, but also be ready for federated learning and MPC in future
  \end{itemize}
\item
  OLAP DBMS

  \begin{itemize}
  \tightlist
  \item
    Often columnar, like Clickhouse and BigQuery
  \item
  \end{itemize}
\end{itemize}

\paragraph{Query \& Processing}\label{query-processing}

\begin{itemize}
\tightlist
\item
  fit in structure of OpenHIE specification
\item
  check which workflows are related to analytics
\item
  Hai calls this `Maintenance'
\end{itemize}

\paragraph{Maintenance}\label{maintenance}

\begin{itemize}
\tightlist
\item
  SQL-on-FHIR Views provide new standard to support mADX aggregate
  reporting !! We need to stress this, because this is an existing
  OpenHIE workflow
\item
  Maintenance-related functions remain the same
\item
  NB: orchestration falls under data provenance
\end{itemize}

\paragraph{Data consumers}\label{data-consumers}

\begin{itemize}
\tightlist
\item
  Many tools, often focus on creating information dashboards and
  visualizations
\item
  Compatibility with processing \& API: which query languages and
  interfaces are supported. Some dialect of SQL, some dialect of NoSQL,
  dataframe API, all of the above?
\end{itemize}

\subsection{Open technologies: deploying Instant OpenHIE with digital
public
goods}\label{open-technologies-deploying-instant-openhie-with-digital-public-goods}

Today, many components of the OpenHIE specification are now available as
a digital public goods, as listed in
Table~\ref{tbl-digital-public-goods}. Typically, these open source
components are intended to support deployments in small countries
(population up to 10 million) or large NGOs out of the box, and should
provide a stepping stone for customized deployments in medium-sized
countries (population around 40 million).\footnote{Although the OpenHIE
  specification does not include details on dimensioning, these are
  typically the requirements that are used within the community. See
  \href{https://wiki.ohie.org/display/CP/Interoperability+Layer+-+Use+Cases+and+Requirements}{OpenHIE
  Community Wiki}.} To further ease the development, configuration and
deployment of health information exchanges, the concept of `Instant
OpenHIE' has been championed to (i) allow implementers to engage with a
preconfigured health information exchange solution and running tools
(based on the architecture) and test their applicability and
functionality with a real health context problem; and (ii) have a
packaged reference version of the OpenHIE architecture that is comprised
of a set of reference technologies and other appropriate tools that form
the building blocks of the health information exchange that can be
configured and extended to support particular use cases
\citep{InstantOpenHIEv2}. Besides the core functional components of the
OpenHIE architecture, the Instant OpenHIE toolkit allows packaging and
integration of generic components such as Identity and Access Managment
(IAM) and a reverse proxy gateway. In the following, we will evaluatie
three of such configurations, with the aim to conceptualize and evaluate
the proposed Data and Analytics Services domain of of the OpenHIE
architecture.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2558}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2558}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4884}}@{}}
\caption{Overview of current open source implementations of components
that fit in the OpenHIE specification. The category Analytics Services
is not a part of the original OpenHIE and is discussed in this paper.
Point-of-Service systems are excluded for brevity.List compiled using
\href{https://wiki.ohie.org/display/DR/OpenHIE+Reference+Technologies}{OpenHIE
Reference Technologies},
\href{https://digitalsquare.org/digital-health-global-goods}{Global
Goods for Digital Health},
\href{https://digitalpublicgoods.net/}{Digital Public Goods Alliance}
and search of open source code repositories. A systematic review of such
digital public goods is beyond the scope of this
document.}\label{tbl-digital-public-goods}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Component
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Digital Public Good
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Component
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Digital Public Good
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Interoperability layer (IOL)} & IOL &
\href{http://openhim.org}{OpenHIM} \\
& & \href{https://www.mhero.org/technology}{mHero} \\
& & \href{https://docs.openfn.org/}{OpenFN} \\
\textbf{Registry Services} & Client Registry (CR) &
\href{https://help.santesuite.org/product-overview/santesuite-products/master-patient-index-santempi}{SanteMPI} \\
& & \href{https://jembi.gitbook.io/jempi/}{JeMPI} \\
& & \href{https://www.openclientregistry.org/}{OpenCR} \\
& & \href{https://www.opencrvs.org/}{OpenCRVS} \\
& Facility Registry (FR) &
\href{https://github.com/intrahealth/gofr}{Global Open Facility Registry
(GOFR)} \\
& & \href{https://geoprismregistry.com/}{GeoPrism Registry (GPR)} \\
& & \href{http://instedd.org/technologies/resource-map/}{Resource
Map} \\
& & \href{https://healthsites.io/}{Healthsite.io} \\
& & \href{https://geoprismregistry.com/}{GeoPrism Registry} \\
& & \href{https://github.com/dhis2}{DHIS2} \\
& Health Worker Registry (HWR) &
\href{https://www.ihris.org/ihris-50}{iHRIS} \\
& Terminology Service (TS) &
\href{https://openconceptlab.org/terminology-service/}{OCL Terminology
Service} \\
& Product Catalog (PC) & \href{https://productcatalog.io/}{PCMT} \\
\textbf{Business Domain Services} & Shared Health Record (SHR) &
\href{https://hapifhir.io/}{HAPI FHIR} \\
& & \href{https://www.health-samurai.io/fhirbase}{Fhirbase} \\
& & \href{https://github.com/microsoft/fhir-server}{FHIR Server for
Azure} \\
& & \href{https://github.com/I-TECH-UW/shared-health-record}{I-TECH-UW
SHR} \\
& Health Management Information System (HMIS)\footnote{Note that there
  is often a confusion on the acronym HMIS. Strictly speaking, the
  OpenHIE specification uses HMIS to refer to a Health Management
  Information System that is part of the Business Domain Services.
  Sometimes HMIS is used to refer to a Hospital Management Systems in
  the Point-of-Service domain, synonymous with an Electronic Medical
  Record (EMR) system.} & \href{http://dhis2.org/}{DHIS2} \\
& Finance and Insurance Service (FIS) &
\href{http://openimis.org/}{OpenIMIS} \\
& Logistics Management Information System (LMIS) &
\href{http://openlmis.org/}{OpenLMIS} \\
& & \href{https://openboxes.com/features/}{OpenBoxes} \\
\textbf{Generic} & Identity and Access Management (IAM) &
\href{https://www.keycloak.org/}{Keycloak} \\
& Gateway \& proxy & \href{https://github.com/google/fhir-gateway}{FHIR
Information Gateway} \\
& Admin dashboard for SHR &
\href{https://github.com/onaio/fhir-web}{OpenSRP FHIR Web} \\
& Configuration and deployment &
\href{https://jembi.gitbook.io/instant-v2/}{Instant OpenHIE} \\
\end{longtable}

\subsection{Open content}\label{open-content}

\begin{itemize}
\tightlist
\item
\end{itemize}

\section{Evaluation}\label{evaluation}

\subsection{Jembi OpenHIM platform}\label{jembi-openhim-platform}

To evaluate the extended OpenHIE architecture described above, we first
consider the OpenHIM Platform. The Open Health Information Mediator
(OpenHIM, \href{OpenHIMhttp://openhim.org/}{http://openhim.org/}))
component is the reference implementation of the Interoperability Layer
(IOL) as defined in the OpenHIE specification. The most current version
(8.4.2 at the time of writing) provides all the core functions including
central point of access for the services of the HIE; routing functions;
central logging for auditing and debugging purposes; and
orchestration/mediation mechanisms to co-ordinate requests. By
extension, the OpenHIM Platform
(\url{https://jembi.gitbook.io/openhim-platform}) is a reference
implementation of a set of Instant OpenHIE configurations, refered to as
`recipes' in the documentation. In the following we will evaluate the
recipe for ``a central data repository with a data warehouse'' that
provides ``A FHIR-based Shared Health record linked to a Master Patient
Index (MPI) for linking and mathing patient demographics and a default
reporting pipeline to transform and visualise FHIR data''
(\url{https://jembi.gitbook.io/openhim-platform/recipes/central-data-repository-with-data-warehousing}).

\begin{figure}

\centering{

\includegraphics{images/openhim-platform.png}

}

\caption{\label{fig-openhim-platform}Overview of the default data stack
of the OpenHIM Platform. The default stack (top, red) consists of Kafka,
Clickhouse and Superset. An alternative solution based on the ELK stack
is also supported (bottom, orange), consisting of Elasticsearch,
Logstash and Kibana.}

\end{figure}%

Figure~\ref{fig-openhim-platform} shows a schematic overview of two data
stacks that are supported in the OpenHIM platform. The Shared Health
Record (SHR, implemented with HAPI FHIR server) and the Client Registry
(CR, implemented with JeMPI server) are the sources that store clinical
FHIR data and patient demographic data, respectively. The default data
stack is based on streaming ingestion using Kafka into a Clickhouse
database. As part of the ingestion, incoming FHIR bundles that contain
multiple FHIR resources are unbundled in separate topics using a generic
Kafka utility component. Subsequently, each FHIR resource topic is
flatted with Kafka mappers that use FHIRPath. Superset is used as the
tool for consuming the data to create dashboard visualizations.

The OpenHIM platform also support data and analytics based on the ELK
stack, where data is ingested in bulk using Logstash, stored in
Elasticsearch and made available for consumption in Kibana. Also here,
the incoming FHIR bundles are unbundled in Logstash into separate FHIR
resources. However, given that Elasticsearch is a document-based search
engine, the FHIR resources are stored as-is with no flattening.
Exploring and analysing the data requires writing queries in
Elasticsearch Query Language (ES\textbar QL), either through the query
interface of Elasticsearch or using Kibana.

Evaluating these two data stacks, we see the following:

\begin{itemize}
\tightlist
\item
  Pattern of flattening FHIR resources with FHIRPath expressions is very
  close to the idea of SQL-on-FHIR. Although it doesn't adhere to this
  new standard in the strict sense, the philosophy of generating tabular
  views is the same
\item
  When using the ELK stack, flattening is done at the end.
  Implementations of FHIRPath support Elasticsearch as an execution
  engine, also here
\item
  Main limitations: both Clickhouse en Elasticsearch don't follow
  decomposition of storage, compute and UI. Therefore, downward
  scaleability is limited.
\end{itemize}

\subsection{ONA OpenSRP 2}\label{ona-opensrp-2}

Continuing our evaluation of the extended OpenHIE architecture, we can
see a different flavor in the implementation driven by Ona. Ona is a
social enterprise that has pioneered the adoption of FHIR data standard
via the development of OpenSRP2, a FHIR-based data collection app built
using Google's FHIR SDK and focused on enabling offline-first workflows
for community-based care. OpenSRP 2 is a complete rewrite of the
original OpenSRP application, a global public good maintained by Ona and
deployed in XX countries worldwide.

OpenSRP2 applications are currently implemented in the field in three
countries (Uganda, Liberia, and Madagascar) in collaboration with local
Ministries of Health and with international donors such as UNICEF,
supporting a variety of different workflows including antenatal care
(ANC), postnatal Care (PNC), immunization, and last-mile logistics.
Besides the OpenSRP Android application and HAPI-FHIR backend, in each
of its projects Ona also implements a companion set of tools that
support analytics and various reporting needs.

\subsubsection{Requirements for data
sharing}\label{requirements-for-data-sharing}

Based on years of work in global health, Ona has learned that the data
stack implemented to support a national-scale implementation of its
FHIR-based application responds to the following requirements.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2857}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7143}}@{}}
\caption{Requirements and rationale for open health data platform
developed and used by ONA, based on OpenSRP 2
(\url{https://opensrp.io/}).}\label{tbl-ona-requirements}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Requirement
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Rationale
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Requirement
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Rationale
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Ingest data from multiple sources, both FHIR and non-FHIR based. & While
most health record data can be collected and aggregated in FHIR,
Ministries of Health rely on other data sources to govern their
operations. For example, operationalizing an immunization campaign
usually includes tracking against specific targets for locations to be
visited on specific days and number of children to immunize per day.
Such targets are often stored in spreadsheets or other applications
where the data is not FHIR. \\
Ingest data in batches. & Most data ingestion can happen in batches,
since Ona's applications are deployed in hard to reach areas where
connectivity is an issue. Data ingestion closer to real-time can be
relevant for disaster-response and other time-sensitive applications,
but this is not a priority. \\
Support national-scale data volumes. & A data store that can grow from
dozens to thousands of devices and where data can be aggregated up to
the national level, matching the scale of implementation of data
collection applications in the field. \\
Pre-compute complex business metrics. & Reporting on health systems
requires pre-computing complex metrics and often performing cohort
analyses to map trends in service provision. For example, understanding
quality of care for children requires computing metrics such as the
percentage of children fully immunized on schedule (i.e.~children 6-59
months that have received the set of vaccines required by the Ministry
of Health, and have received each of those vaccines within the expected
age-window). For Business Intelligence applications, calculating such a
vital metric cannot be performed at run time, to avoid long and
expensive queries. \\
Outbound integrations. & While aggregated data and reports should be
accessible by other applications such as BI platforms via pulls, there
should be an easy integration framework to push data to other
applications used by the Ministry of Health for other purposes, such as
DHIS2 for health systems management or RapidPro for communications with
program beneficiaries. \\
Open source and easily deployable in-country. & Given the extremely
sensitive nature of health data, it is paramount for governments to have
the flexibility to deploy the stack in various different environments,
both on premise and in private clouds. \\
\end{longtable}

The architecture Learning from experience in the field and internal
research and development, Ona has developed preferences for a specific
data stack responding to the aforementioned requirements.

{[}{[}graphic{]}{]}

Core toolings in the stack include: Data ingestion with Airbyte. Ona
uses Airbyte as the primary data ingestion tool, leveraging the wide
array of connectors that come standard with the application as well as a
dedicated suite of connectors developed internally by Ona, including
HAPI FHIR, RapidPro, Ona Data, Kobo Toolbox and others.\\
Data storage with Clickhouse. While different health projects have
varying requirements, Ona has found success in using Clickhouse as the
main analytics data store in its most recent implementations. Clickhouse
supports the scale required for analytics at a national level, as well
as the speed that enables cross-application integrations and more
real-time analytics. For example, in Madagascar Ona uses its reporting
suite to identify facilities with stock in need of maintenance and can
trigger the scheduling of a maintenance visit ad hoc. Data
transformation with dbt. Following global best practice, Ona leverages
dbt to segregate the data warehouse in different levels (staging, marts,
metrics), as well as pre-computing complex indicators for ease of
reporting and for transmissions into other systems. For example, in
Liberia Ona implements OpenSRP at community health worker level, but can
aggregate immunization data at facility level in the data warehouse and
then push quarterly summary metrics to DHIS2. No recommendation on
reporting / BI tooling. Ona recognizes that business users have their
own strong preferences for BI tooling, and some already have licenses
for specific software, so the architecture is flexible to provide easy
connections to different BI tools.

Evaluation

Evaluating the data stack, we see the following: Use of generic
best-of-breed tooling. Ona focused on utilizing Open HIE tools that are
widely adopted outside of the global health and development sectors.
This approach aims to provide assurance on two main fronts, the ability
to handle performance at scale and the long term dependability of the
tools, rather than relying on smaller projects with uncertain long term
funding or unproven implementations. Columnar data warehouse for
analytics. The scale of Ona's project requires the implementation of a
dedicated database for analytics. While original data can still be
stored as parquet or other file system, being able to ingest it into a
relational data store allows to create well defined indicators. Using
clickhouse as a tool helps and combine the need accuracy with the speed
of reporting as new data is ingested.\\
Strong emphasis on SQL. While Ona has tested and experimented with
FHIR-specific tooling, such as the definition of data projections using
sql-on-fhir, Ona found that relying on sql for coding business logic
remained the faster and most scalable approach.

In summary, for Ona building analytics with FHIR data looks similar to
building analytics with any other type of data. While FHIR provides a
clear and standard data model, managing information for most health
systems requires custom integration of data between different sources,
as well as computing indicators using business logic specific to the
needs of the local users. Building upon well established best-of-breed
tools allows Ona to implement FHIR applications at scale and provide
trusted analytics on top.

\subsection{PharmAccess demonstrator Momcare
programme}\label{pharmaccess-demonstrator-momcare-programme}

MomCare was launched in Kenya
\citep{huisman2022digital, sanctis2022maintaining} and Tanzania
\citep{shija2021access, mrema2021application} in 2017 and 2019
respectively, with the objective to improve health outcomes for maternal
and antenatal care. MomCare distinguishes two user groups: mothers are
supported during their pregnancy through reminders and surveys, using
SMS as the digital mode of engagement. Health workers are equipped with
an Android-based application, in which visits, care activities and
clinical observations are recorded. Reimbursements of the maternal
clinic are based on the data captured with SMS and the app, thereby
creating a conditional payment scheme, where providers are partially
reimbursed up-front for a fixed bundle of activities, supplemented by
bonus payments based on a predefined set of care activities.

In its original form, the MomCare programme used closed digital
platforms. In Kenya, M-TIBA is the primary digital platform, on top of
which a relatively lightweight custom app has been built as the
engagement layer for the health workers \citep{huisman2022digital}.
M-TIBA provides data access through its data warehouse platform for the
MomCare programme, however, this is not a standardized, general purpose
API. In the case of Tanzania, a stand-alone custom app is used which
does not provide an interface of any kind for interacting with the
platform \citep{mrema2021application}. Given these constraints, the
first iteration of the MomCare programme used a custom-built data
warehouse environment as its main data platform, on which data
extractions, transformations and analysis are performed to generate the
operational reports. Feedback reports for the health workers, in the
form of operational dashboards, are made accessible through the app.
Similar reports are provided to the back-office for the periodic
reimbursement to the clinics.

Clearly, a more open and scaleable platform was required if MomCare was
to be implemented in more regions. This need led to a redesign of the
underlying technical infrastructure of the MomCare project. The
objectives of this work were in fact to demonstrate a solution design
that could support the first three types of data sharing. First, to
investigate the viability of using FHIR for bulk data sharing, MomCare
Tanzania was used a testbed to assess the complexity and effort required
to implement the facade pattern to integrate the legacy system into the
FHIR data standard. Using the longitudinal dataset from approximately 28
thousand patient records, FHIR transformations script were developed and
deployed using the mediator function of the IOL. The data was
transformed into 10 FHIR v4 resources and the conceptual data model of
the existing MomCare app could readily be transformed into the FHIR
standard using SQL and validated with a Python library
\citep{islam2023fhir}. The largest challenge during the transformation
process pertained to the absence of unique business identifiers for
patients and healthcare organizations. For patients, either the mobile
phone number or the healthcare insurance number was taken, depending on
availability. A combination of name, address and latitude/longitude
coordinates were used to uniquely identify organizations and locations,
as Tanzania does not have a system in place for this purpose.

The second objective was to reproduce existing analytic reports, using
the bulk FHIR data format as input. Here, the focus was to standardize
the logic required for producing metrics and reports. The transformed
and validated data is uploaded into the FHIR server on a daily basis
using an automated cloud function. Analysis of bulk data was done by
directly reading the standard newline delimited JSON into the Python
pandas data analysis library. Cross checking the output with queries on
the original data confirmed that the whole data pipeline produced
consistent results. For example, the report of the antenatal coverage
metric (number of pregnancies with four or more visits) could be
reproduced per patient journey and aggregated (per year, per
organization etc.) as required for the MomCare reports.

TO DO: explain logic of patient-timeline table. Write standard
transformation to go from FHIR resources to this standard table. On top
of that the actual metrics and reporting. Explain serverless: we wanted
to get rid of resource-heavy data visualization tools. This led to the
idea of serverless: using duckdb-wasm and pipelines of cloud functions.

The third objective was to run a technical feasibility test for
federated analytics. Using the MPC platform of Roseman Labs, we managed
to do aggregations in the blind \ldots{} TO DO: explain that we managed
to reproduce the reports we generated in the clear, but then in the
blind. Note, however, that in the remainder we will focus on first two
types of data sharing.

Based on these experiments, we arrived at the following design for the
data \& analytics services

\begin{itemize}
\tightlist
\item
  Use `serverless' file-based storage: bulk copy of data as-is in
  parquet

  \begin{itemize}
  \tightlist
  \item
    Tension: how to manage change data capture
  \item
    Tension: how to manage access rights
  \end{itemize}
\item
  Use SQL-on-FHIR-v2 to create tabular views.

  \begin{itemize}
  \tightlist
  \item
    Example: patient timeline
  \item
    TO DO: rewrite patient timeline queries with SQL-on-FHIR-v2 and run
    it with Pathling
  \end{itemize}
\item
  Use semantic modeling layer to define metrics

  \begin{itemize}
  \tightlist
  \item
    There are many options: dbt, cube.dev
  \item
    Fulfills same function as ADX/mADX IHE profile in OpenHIE
    specification
  \item
    Tension: going from patient-timeline to reported metrics still isn't
    standardized. This is where Ibis/Substrait comes in. Substrait as IR
    for cross-language serialization for relational algebra. Can be
    executed on different backends. Write once, run on different
    engines.
  \end{itemize}
\item
  Distribute and publish reports on resource-constrained devices

  \begin{itemize}
  \tightlist
  \item
    duckdb
  \item
    sveltekit
  \end{itemize}
\end{itemize}

TO DO: Add diagram

\subsubsection{Level of opennness}\label{level-of-opennness}

TO DO: evaluate openness of OpenHIM platform

\subsubsection{Core maintenance
functions}\label{core-maintenance-functions}

TO DO: evaluate maintenance functions

\subsubsection{Downward scaleability}\label{downward-scaleability}

TO DO: evaluate downward scalability

\section{Discussion}\label{discussion}

\subsection{Openness of data
platforms}\label{openness-of-data-platforms}

We specifically address the notion of openness of OHDPs in LMICs in
terms of the design-related questions put forward by de Reuver at al.11:

\begin{itemize}
\tightlist
\item
  Object of openness: what data-related resources should data platforms
  make available when opening up (e.g.~data, data products, datadriven
  insights, analytics modules)? Which user groups derive value from
  accessing data-related resources from data platforms (e.g.~data
  providers, data users, intermediaries, developers)?
\item
  Unit of analysis: what is platform-to-platform openness in the context
  of data platforms, given the expectation that different OHDPs will
  emerge at various aggregation levels? How do we distinguish
  meta-platforms, forking, and platform interoperability?
\item
  Risk of openness: What are the novel (negative) implications of
  opening up data platforms? How can reflexivity in design help
  providers to resolve the negative implications of openness?
\item
  Answers/insights to above:

  \begin{itemize}
  \tightlist
  \item
    Openness of standardized view on FHIR data and cross-language
    serialization of relational algebra makes it possible to fully
    standardize the workflow from start to finish
  \item
    Platform-to-platform: MPC
  \item
    Risk of openness: difficult to answer \ldots{}
  \end{itemize}
\end{itemize}

\subsection{Comparison with HMIS
component}\label{comparison-with-hmis-component}

\begin{itemize}
\tightlist
\item
  Workflow requirements: Report aggregate data (link): receiver is HMIS,
  mADX
\item
  Functional requirements:
  https://guides.ohie.org/arch-spec/openhie-component-specifications-1/openhie-health-management-information-system-hmis
\end{itemize}

Requirements are similar, but implementation differs: Datamodel is
non-FHIR, focused on DataValue, which conceptually equates to FHIR
Measure

\subsection{The need to a semantic
layer?}\label{the-need-to-a-semantic-layer}

\begin{itemize}
\tightlist
\item
  FHIR and FAIR

  \begin{itemize}
  \tightlist
  \item
    How does FHIR relate to approaches taken by the FAIR community,
    which tend to take more an approach of using knowledge graphs. For
    example, VODAN Africa
    \citep{gebreslassie2023fhir4fair, purnamajati2022data}.
  \item
    FAIR principles vs FHIR graph: is FHIR a FAIR Data Object
  \end{itemize}
\item
  Since we use FHIR, we don't need a semantic layer because that is
  already provided
\item
  We do need different semantic layer, namely with metrics. Explain
  different types of semantics.

  \begin{itemize}
  \tightlist
  \item
    The metrics layer same function as CQL. Discuss CQL vs generic
    metrics layer.
  \end{itemize}
\end{itemize}

\subsection{Attribute-based access
control}\label{attribute-based-access-control}

\begin{itemize}
\tightlist
\item
  TO DO: if you have generated flattened SQL tables, how are you going
  to manage security?
\item
  Cerbos, attribute based on lineage or anonymized tables
\item
  Catalogs solve this: Tabular.io, Google BigLake. What is open source
  option?
\end{itemize}

\subsection{Federated learning and multiparty
computation}\label{federated-learning-and-multiparty-computation}

\begin{itemize}
\tightlist
\item
  data stations??!
\end{itemize}

\section{Abbreviations}\label{abbreviations}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
ACID & Atomicity, Consistency, Isolation, and Durability \\
CLI & Command-line Interface \\
CR & Client Registry \\
ELK & Elasticsearch, Logstach and Kibana stack \\
ELT & Extract, Load and Transform \\
FAIR & Findable, Accessible, Interoperable and Reusable \\
FHIR & Fast Healthcare Interoperability Resources \\
FL & Federated learning \\
OHDP & Health data platform, explicitly differentiated from health
digital platform \\
HIE & Health Information Exchange \\
IR & Intermediate Representation \\
LMIC & Low- and middle income countries \\
MPC & Multiparty Computation \\
PET & Privacy-enhancing technologies \\
SHR & Shared Health Record \\
\end{longtable}

  \bibliography{pharmaccess.bib}

\end{document}