### 2024-08-26 index.qmd
wip

---
title: "Design and implemention of open health data systems with standardized data and analytics capabilities for low- and middle income countries"
format:
  elsevier-html:
    toc: true
    toc-depth: 3
    fig-cap-location: top
    code-fold: true
  pdf: default
    # keep-tex: true
    # journal:
    #   name: JIMR
    #   cite-style: authoryear
author:
  - name: Daniel Kapitan
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands
      - name: Eindhoven University of Technology
        city: Eindhoven
        country: the Netherlands
    orcid: 0000-0001-8979-9194
    email: daniel@kapitan.net
  - name: Femke Heddema
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands
  - name: Julie Fleischer
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands
  - name: Irni Gemzon
    affiliations:
      - name: PharmAccess Foundation
        city: Amsterdam
        country: the Netherlands
  - name: Chris Ihure
    affiliations:
      - name: PharmAccess Kenya
        city: Nairobi
        country: Kenya
  - name: Steven Wanyee
    affiliations:
      - name: IntelliSOFT
        city: Nairobi
        country: Kenya
  - name: Alessandro Pietrobon 
    affiliations:
      - name: ONA
        city: Nairobi
        country: Kenya
  - name: Ryan Chrichton
    affiliations:
      - name: Jembi Health Systems
        city: Durban
        country: South Africa
  - name: John Grimes
    affiliations:
      - name: Australian e-Health Research Centre
        city: Brisbane
        country: Australia
  # - name: Paula van Brakel
  #   affiliations:
  #     - name: PharmAccess Foundation
  #       city: Amsterdam
  #       country: the Netherlands
  # - name: Mark van der Graaf
  #   affiliations:
  #     - name: PharmAccess Foundation
  #       city: Amsterdam
  #       country: the Netherlands
  # - name: Nicole Spieker
  #   affiliations:
  #     - name: PharmAccess Foundation
  #       city: Amsterdam
  #       country: the Netherlands

abstract: |
  TO DO: add abstract.
keywords: [Analytics-on-FHIR, SQL-on-FHIR, HIE, data platforms, LMICs, digital health]
reference-section-title: References
bibliography: pharmaccess.bib
---

## Introduction {#sec-introduction}

### The global digital health ecosystem is still project-centric

Digital transformation have the potential to make inclusive health systems in low- and middle income countries (LMICs) a reality by connecting patients and professionals, using data as a currency of trust. Digital technologies have the potential to increase the availability, accessibility, acceptability, and quality of health services as we know them; change nature and direction of health care, and the possibility of making it more preventive, personalised, and mobile through the use of such technologies; and enfranchise patients and communities (and particularly vulnerable groups including children and young people) and evolve their relationship with health professionals and providers, thus helping shape the health system according to the needs of the patients and communities [@kickbusch2021lancet].

Yet, to date the global digital health ecosystem is still project-centric, resulting in data fragmentation and technology lock-in, compromising health care delivery [@mehl2023fullstac]. There is no dearth of the number of digital health interventions: sub-Saharan Africa has seen over 700 such projects in the past 10 years. There is, however, a fundamental lack in coordination, integration, scalability, sustainability, and equitable distribution of investments in digital health. Health policymakers and the global health community at large need to urgently institute coordination mechanisms to terminate unending duplication and disjointed vertical implementations and manage solutions for scale [@karamagi2022ehealth].

To achieve scaleability of digital health interventions, we need to design and implement standardized open health data systems (OHDS) and their associated ecosystems that can support the wider health sector in improving along five dimensions, namely [@kelley2020digital]:

1. overall quality and continuity of care;
2. adherence to clinical guidelines and best practices;
3. efficiency and affordability of services and health commodities, by reducing duplication of effort and ensuring effective use of time and resources;
4. health-financing models and processes, regulation, oversight, and patient safety resulting from increased availability of performance data and reductions in errors; and
5. health policy-making and resource allocation based on better quality data .

### Data & analytics functions are essential in open health data systems

Learnings from digital health interventions have shown that improvements along the five aforementioned dimensions is indeed within reach. As an example, in our experience with the MomCare programme, we have demonstrated that routinely collected health data, combined with financial data, can be effectively used to gain insight into the continuity of care, improve clinical adherence whilst maintaining efficiency and affordability of health services in LMICs. By actively coaxing pregnant mothers to undergo 4 antenatal check-ups, outcomes were improved whilst maintaining the average cost of MCH services [@huisman2022digital;@sanctis2022maintaining;@izudi2023experiences].

MomCare, like most other digital health interventions, is critically dependent on the availability of data and analytics functions within the OHDS. For example, the widely used DHIS2 reports that are most commonly used in LMICs contained insufficient information to run the programme. More detailed analytics at the patient level were required. It is, in fact, a widely held belief that data & analytics are an essential part of OHDSs. This is exemplified how the OpenHIE reference architecture [@openhie] has been adopted by many sub-Saharan African countries as the blueprint for implementing nation-wide health information exchanges (HIE) [@mamuye2022health], including Nigeria [@dalhatu2023paper], Kenya [@mbugua2021adoption] and Tanzania [@nsaghurwe2021one]. These countries have, as a matter of course, extended the framework to include “data & analytics services” as an additional domain [@mbugua2021adoption;@dalhatu2023paper]. The rationale for this addition is to facilitate secondary reuse of health data for academic research, real-world evidence studies etc. which can be framed within the context of ongoing efforts towards Findable, Accessible, Interoperable and Reusable (FAIR) sharing of health data [@guillot2023fair].

## Extending OpenHIE to include modern data and analytics standards

If we intend to use the OpenHIE standard not only for primary health data sharing, but also to support data and analytics for secondary health data sharing, we need to extend its specification. The main contribution of this paper is to propose how recent standards and open source implementations from the data engineering community could be integrated into the OpenHIE framework, as illustrated in @fig-ohie.

![Proposed extension of the OpenHIE architecture that includes "Data and Analytics Services" as an additional service domain.](./images/openhie-extended-architecture.png){#fig-ohie}

The original OpenHIE specification discerns four domains, namely Point-of-Service systems, the Interoperability Layer, Common Services and Business Services. We propose to extend the OpenHIE architecture with a "Data and Analytics Services" domain with different zones taken from the data lakehouse architecture, which currently is the most commonly used design pattern in this domain [@armbrust2021lakehouse;@hai2023data;@harby2022data;@harby2024data]. Lakehouses typically have a zonal architecture that follow the Extract-Load-Transform pattern (ELT) where data is ingested from the source systems in bulk (E), delivered to storage with aligned schemas (L) and transformed into a format ready for analysis (T) [@hai2023data]. The discerning characteristic of the lakehouse architecture is its foundation on low-cost and directly-accessible storage that also provides traditional database management and performance features such as ACID transactions, data versioning, auditing, indexing, caching, and query optimization [@armbrust2021lakehouse]. Lakehouses thus combine the key benefits of data lakes and data warehouses: low-cost storage in an open format accessible by a variety of systems from the former, and powerful management and optimization features from the latter.

Following the terminology proposed by Hai et al. [@hai2023data], @tbl-data-and-analysis-services summarizes how the different zones of the lakehouse architecture can be adapted for healthcare and integrated into the OpeHIE specification. Taking Fast Healthcare Interoperability Resources (FHIR) as the open data standard, we envisage the extended OpenHIE architecture to include a 'FHIR lakehouse'[^1].

[^1]: We have argued the choice of FHIR as the common data model elsewhere, working paper to be submitted ([link](https://plugin-healthcare.github.io/viewpoint-jimr/)).

+-------------------+---------------------------------------------------------------------------------------+
| Service           | Functional requirements                                                               |
+===================+=======================================================================================+
| Ingestion         | - Bulk                                                                                |
|                   | - Streaming                                                                           |
+-------------------+---------------------------------------------------------------------------------------+
| Storage           | - File-based blob storage                                                             |
|                   | - Database optimized for online analytical processing (OLAP)
+-------------------+---------------------------------------------------------------------------------------+
| Maintenance       | - SQL-on-FHIR View defintions                                                         |
|                   | - Catalog and other maintenance-related functions as defined by Hai et al.            |
+-------------------+---------------------------------------------------------------------------------------+
| Processing        | - SQL-on-FHIR Runner                                                                  |
| & API             | - Execution engine on tabular data as defined in composable data stack                |
|                   | - Capability to participate as a node in federated learning / MPC network             |
|                   | - Read-only access to storage                                                         |  
+-------------------+---------------------------------------------------------------------------------------+
| Data              | - SQL interactive development environment (IDE)                                       |
| Consumer          | - Interactive notebook computing environment [@granger2021jupyter]                    |
|                   | - BI reporting, dashboarding and data visualization                                   |  
+-------------------+---------------------------------------------------------------------------------------+
: Adaptation of lakehouse zones for healthcare specific standardization of Data and Analysis Services {#tbl-data-and-analysis-services}

[TO DO: need to explain table (is too condensed now) and especially FHIR specific components.]

In the remainder of this paper, we first describe in more detail how the FHIR lakehouse pattern can be implemented using existing open source technologies. We do so by presenting work done at PharmAccess. Code and digital artifacts are made available [TO DO: include links to support GitHub repositories].

Subquently, we compare this solution design with two widely used OpenHIE-compliant frameworks, namely the OpenHIM platform ([https://jembi.gitbook.io/openhim-platform/](https://jembi.gitbook.io/openhim-platform/)) and the OnaData platform [https://ona.io/home/products/ona-data/features/](https://ona.io/home/products/ona-data/features/). We take a narrative approach in presenting our design and the case studies, surveying existing scientific studies on OHDSs, focusing on the seminal reports and subsequently searching forward citations. In addition, we have searched the open source repositories (most notably GitHub) and the online communities (OpenHIE community, FHIR community) to search for relevant open standards, technologies and architectures. This paper should not be considered as a proper systematic review.

Finally, we discuss our findings and propose routes for future development.

The main contributions of this paper are i) description of a framework for the components of the Data & Analysis Services that builds on current best practices from the data engineering community into the OpenHIE framework; ii) evaluation of different implementations and design options for type 1 and 2 data sharing within an extended OpenHIE architecture; and iii) provide open content for the MomCare implementation to facilitate adaptation and deployment. As such, it aims to inform future developments and implementation of open digital health platforms in LMICs.

## Case study of implementations

### PharmAccess demonstrator Momcare programme

MomCare was launched in Kenya [@huisman2022digital;@sanctis2022maintaining] and Tanzania [@shija2021access;@mrema2021application] in 2017 and 2019 respectively, with the objective to improve health outcomes for maternal and antenatal care. MomCare distinguishes two user groups: mothers are supported during their pregnancy through reminders and surveys, using SMS as the digital mode of engagement. Health workers are equipped with an Android-based application, in which visits, care activities and clinical observations are recorded. Reimbursements of the maternal clinic are based on the data captured with SMS and the app, thereby creating a conditional payment scheme, where providers are partially reimbursed up-front for a fixed bundle of activities, supplemented by bonus payments based on a predefined set of care activities.

In its original form, the MomCare programme used closed digital platforms. In Kenya, M-TIBA is the primary digital platform, on top of which a relatively lightweight custom app has been built as the engagement layer for the health workers [@huisman2022digital]. M-TIBA provides data access through its data warehouse platform for the MomCare programme, however, this is not a standardized, general purpose API. In the case of Tanzania, a stand-alone custom app is used which does not provide an interface of any kind for interacting with the platform [@mrema2021application]. Given these constraints, the first iteration of the MomCare programme used a custom-built data warehouse environment as its main data platform, on which data extractions, transformations and analysis are performed to generate the operational reports. Feedback reports for the health workers, in the form of operational dashboards, are made accessible through the app. Similar reports are provided to the back-office for the periodic reimbursement to the clinics.

Clearly, a more open and scaleable platform was required if MomCare was to be implemented in more regions. This need led to a redesign of the underlying technical infrastructure of the MomCare project. The objectives of this work were in fact to demonstrate a solution design that could support the first three types of data sharing. First, to investigate the viability of using FHIR for bulk data sharing, MomCare Tanzania was used a testbed to assess the complexity and effort required to implement the facade pattern to integrate the legacy system into the FHIR data standard. Using the longitudinal dataset from approximately 28 thousand patient records, FHIR transformations script were developed and deployed using the mediator function of the IOL. The data was transformed into 10 FHIR v4 resources and the conceptual data model of the existing MomCare app could readily be transformed into the FHIR standard using SQL and validated with a Python library [@islam2023fhir]. The largest challenge during the transformation process pertained to the absence of unique business identifiers for patients and healthcare organizations. For patients, either the mobile phone number or the healthcare insurance number was taken, depending on availability. A combination of name, address and latitude/longitude coordinates were used to uniquely identify organizations and locations, as Tanzania does not have a system in place for this purpose.

The second objective was to reproduce existing analytic reports, using the bulk FHIR data format as input. Here, the focus was to standardize the logic required for producing metrics and reports. The transformed and validated data is uploaded into the FHIR server on a daily basis using an automated cloud function. Analysis of bulk data was done by directly reading the standard newline delimited JSON into the Python pandas data analysis library. Cross checking the output with queries on the original data confirmed that the whole data pipeline produced consistent results. For example, the report of the antenatal coverage metric (number of pregnancies with four or more visits) could be reproduced per patient journey and aggregated (per year, per organization etc.) as required for the MomCare reports.

TO DO: explain logic of patient-timeline table. Write standard transformation to go from FHIR resources to this standard table. On top of that the actual metrics and reporting. Explain serverless: we wanted to get rid of resource-heavy data visualization tools. This led to the idea of serverless: using duckdb-wasm and pipelines of cloud functions.

The third objective was to run a technical feasibility test for federated analytics. Using the MPC platform of Roseman Labs, we managed to do aggregations in the blind ...
TO DO: explain that we managed to reproduce the reports we generated in the clear, but then in the blind. Note, however, that in the remainder we will focus on first two types of data sharing.

Based on these experiments, we arrived at the following design for the data & analytics services

- Use 'serverless' file-based storage: bulk copy of data as-is in parquet
  - Tension: how to manage change data capture
  - Tension: how to manage access rights
- Use SQL-on-FHIR-v2 to create tabular views.
  - Example: patient timeline
  - TO DO: rewrite patient timeline queries with SQL-on-FHIR-v2 and run it with Pathling
- Use semantic modeling layer to define metrics
  - There are many options: dbt, cube.dev
  - Fulfills same function as ADX/mADX IHE profile in OpenHIE specification
  - Tension: going from patient-timeline to reported metrics still isn't standardized. This is where Ibis/Substrait comes in. Substrait as IR for cross-language serialization for relational algebra. Can be executed on different backends. Write once, run on different engines. 
- Distribute and publish reports on resource-constrained devices 
  - duckdb
  - sveltekit

TO DO: Add diagram

### Jembi OpenHIM platform

To evaluate the extended OpenHIE architecture described above, we first consider the OpenHIM Platform. The Open Health Information Mediator (OpenHIM, [http://openhim.org/](OpenHIMhttp://openhim.org/))) component is the reference implementation of the Interoperability Layer (IOL) as defined in the OpenHIE specification. The most current version (8.4.2 at the time of writing) provides all the core functions including central point of access for the services of the HIE; routing functions; central logging for auditing and debugging purposes; and orchestration/mediation mechanisms to co-ordinate requests. By extension, the OpenHIM Platform ([https://jembi.gitbook.io/openhim-platform](https://jembi.gitbook.io/openhim-platform)) is a reference implementation of a set of Instant OpenHIE configurations, refered to as 'recipes' in the documentation. In the following we will evaluate the recipe for "a central data repository with a data warehouse" that provides "A FHIR-based Shared Health record linked to a Master Patient Index (MPI) for linking and mathing patient demographics and a default reporting pipeline to transform and visualise FHIR data" ([https://jembi.gitbook.io/openhim-platform/recipes/central-data-repository-with-data-warehousing](https://jembi.gitbook.io/openhim-platform/recipes/central-data-repository-with-data-warehousing)).

![Overview of the default data stack of the OpenHIM Platform. The default stack (top, red) consists of Kafka, Clickhouse and Superset. An alternative solution based on the ELK stack is also supported (bottom, orange), consisting of Elasticsearch, Logstash and Kibana.](images/openhim-platform.png){#fig-openhim-platform}

@fig-openhim-platform shows a schematic overview of two data stacks that are supported in the OpenHIM platform. The Shared Health Record (SHR, implemented with HAPI FHIR server) and the Client Registry (CR, implemented with JeMPI server) are the sources that store clinical FHIR data and patient demographic data, respectively. The default data stack is based on streaming ingestion using Kafka into a Clickhouse database. As part of the ingestion, incoming FHIR bundles that contain multiple FHIR resources are unbundled in separate topics using a generic Kafka utility component. Subsequently, each FHIR resource topic is flatted with Kafka mappers that use FHIRPath. Superset is used as the tool for consuming the data to create dashboard visualizations.

The OpenHIM platform also support data and analytics based on the ELK stack, where data is ingested in bulk using Logstash, stored in Elasticsearch and made available for consumption in Kibana. Also here, the incoming FHIR bundles are unbundled in Logstash into separate FHIR resources. However, given that Elasticsearch is a document-based search engine, the FHIR resources are stored as-is with no flattening. Exploring and analysing the data requires writing queries in Elasticsearch Query Language (ES|QL), either through the query interface of Elasticsearch or using Kibana.

Evaluating these two data stacks, we see the following:

- Pattern of flattening FHIR resources with FHIRPath expressions is very close to the idea of SQL-on-FHIR. Although it doesn't adhere to this new standard in the strict sense, the philosophy of generating tabular views is the same
- When using the ELK stack, flattening is done at the end. Implementations of FHIRPath support Elasticsearch as an execution engine, also here 
- Main limitations: both Clickhouse en Elasticsearch don't follow decomposition of storage, compute and UI. Therefore, downward scaleability is limited.

### ONA OpenSRP 2

Continuing our evaluation of the extended OpenHIE architecture, we can see a different flavor in the implementation driven by Ona. Ona is a social enterprise that has pioneered the adoption of FHIR data standard via the development of OpenSRP2, a FHIR-based data collection app built using Google’s FHIR SDK and focused on enabling offline-first workflows for community-based care. OpenSRP 2 is a complete rewrite of the original OpenSRP application, a global public good maintained by Ona and deployed in XX countries worldwide.

OpenSRP2 applications are currently implemented in the field in three countries (Uganda, Liberia, and Madagascar) in collaboration with local Ministries of Health and with international donors such as UNICEF, supporting a variety of different workflows including antenatal care (ANC), postnatal Care (PNC), immunization, and last-mile logistics. Besides the OpenSRP Android application and HAPI-FHIR backend, in each of its projects Ona also implements a companion set of tools that support analytics and various reporting needs.

##### Requirements for data sharing

Based on years of work in global health, Ona has learned that the data stack implemented to support a national-scale implementation of its FHIR-based application responds to the following requirements.

| Requirement | Rationale |
|:-----|:--------------|
| Ingest data from multiple sources, both FHIR and non-FHIR based. | While most health record data can be collected and aggregated in FHIR, Ministries of Health rely on other data sources to govern their operations. For example, operationalizing an immunization campaign usually includes tracking against specific targets for locations to be visited on specific days and number of children to immunize per day. Such targets are often stored in spreadsheets or other applications where the data is not FHIR. |
| Ingest data in batches. | Most data ingestion can happen in batches, since Ona’s applications are deployed in hard to reach areas where connectivity is an issue. Data ingestion closer to real-time can be  relevant for disaster-response and other time-sensitive applications, but this is not a priority.|
| Support national-scale data volumes. | A data store that can grow from dozens to thousands of devices and where data can be aggregated up to the national level, matching the scale of implementation of data collection applications in the field.|
| Pre-compute complex business metrics. | Reporting on health systems requires pre-computing complex metrics and often performing cohort analyses to map trends in service provision. For example, understanding quality of care for children requires computing metrics such as the percentage of children fully immunized on schedule (i.e. children 6-59 months that have received the set of vaccines required by the Ministry of Health, and have received each of those vaccines within the expected age-window). For Business Intelligence applications, calculating such a vital metric cannot be performed at run time, to avoid long and expensive queries.|
| Outbound integrations. | While aggregated data and reports should be accessible by other applications such as BI platforms via pulls, there should be an easy integration framework to push data to other applications used by the Ministry of Health for other purposes, such as DHIS2 for health systems management or RapidPro for communications with program beneficiaries.|
| Open source and easily deployable in-country.| Given the extremely sensitive nature of health data, it is paramount for governments to have the flexibility to deploy the stack in various different environments, both on premise and in private clouds.|

: Requirements and rationale for open health data platform developed and used by ONA, based on OpenSRP 2 ([https://opensrp.io/](https://opensrp.io/)). {#tbl-ona-requirements .striped .hover .responsive}

The architecture
Learning from experience in the field and internal research and development, Ona has developed preferences for a specific data stack responding to the aforementioned requirements.

[[graphic]]

Core toolings in the stack include: 
Data ingestion with Airbyte. Ona uses Airbyte as the primary data ingestion tool, leveraging the wide array of connectors that come standard with the application as well as a dedicated suite of connectors developed internally by Ona, including HAPI FHIR, RapidPro, Ona Data, Kobo Toolbox and others.  
Data storage with Clickhouse. While different health projects have varying requirements, Ona has found success in using Clickhouse as the main analytics data store in its most recent implementations.  Clickhouse supports the scale required for analytics at a national level, as well as the speed that enables cross-application integrations and more real-time analytics. For example, in Madagascar Ona uses its reporting suite to identify facilities with stock in need of maintenance and can trigger the scheduling of a maintenance visit ad hoc. 
Data transformation with dbt. Following global best practice, Ona leverages dbt to segregate the data warehouse in different levels (staging, marts, metrics), as well as pre-computing complex indicators for ease of reporting and for transmissions into other systems. For example, in Liberia Ona implements OpenSRP at community health worker level, but can aggregate immunization data at facility level in the data warehouse and then push quarterly summary metrics to DHIS2.
No recommendation on reporting / BI tooling. Ona recognizes that business users have their own strong preferences for BI tooling, and some already have licenses for specific software, so the architecture is flexible to provide easy connections to different BI tools.

Evaluation

Evaluating the data stack, we see the following:
Use of generic best-of-breed tooling. Ona focused on utilizing Open HIE tools that are widely adopted outside of the global health and development sectors. This approach aims to provide assurance on two main fronts, the ability to handle performance at scale and the long term dependability of the tools, rather than relying on smaller projects with uncertain long term funding or unproven implementations.
Columnar data warehouse for analytics. The scale of Ona’s project requires the implementation of a dedicated database for analytics. While original data can still be stored as parquet or other file system, being able to ingest it into a relational data store allows to create well defined indicators. Using clickhouse as a tool helps and combine the need accuracy with the speed of reporting as new data is ingested.  
Strong emphasis on SQL. While Ona has tested and experimented with FHIR-specific tooling, such as the definition of data projections using sql-on-fhir, Ona found that relying on sql for coding business logic remained the faster and most scalable approach.

In summary, for Ona building analytics with FHIR data looks similar to building analytics with any other type of data. While FHIR provides a clear and standard data model, managing information for most health systems requires custom integration of data between different sources, as well as computing indicators using business logic specific to the needs of the local users. Building upon well established best-of-breed tools allows Ona to implement FHIR applications at scale and provide trusted analytics on top.

### Open technologies: deploying Instant OpenHIE with digital public goods

Today, many components of the OpenHIE specification are now available as a digital public goods, as listed in @tbl-digital-public-goods. Typically, these open source components are intended to support deployments in small countries (population up to 10 million) or large NGOs out of the box, and should provide a stepping stone for customized deployments in medium-sized countries (population around 40 million).[^2] To further ease the development, configuration and deployment of health information exchanges, the concept of 'Instant OpenHIE' has been championed to (i) allow implementers to engage with a preconfigured health information exchange solution and running tools (based on the architecture) and test their applicability and functionality with a real health context problem; and (ii) have a packaged reference version of the OpenHIE architecture that is comprised of a set of reference technologies and other appropriate tools that form the building blocks of the health information exchange that can be configured and extended to support particular use cases [@InstantOpenHIEv2]. Besides the core functional components of the OpenHIE architecture, the Instant OpenHIE toolkit allows packaging and integration of generic components such as Identity and Access Managment (IAM) and a reverse proxy gateway. In the following, we will evaluatie three of such configurations, with the aim to conceptualize and evaluate the proposed Data and Analytics Services domain of of the OpenHIE architecture.

[^2]: Although the OpenHIE specification does not include details on dimensioning, these are typically the requirements that are used within the community. See [OpenHIE Community Wiki](https://wiki.ohie.org/display/CP/Interoperability+Layer+-+Use+Cases+and+Requirements).

We also posit that a decentralized platform is more conducive to realize a solidarity-based approach to health data sharing that i) gives people a greater control over their data as active decision makers; ii) ensures that the value of data is harnessed for public good; and iii) moves society towards equity and justice by counteracting dynamics of data extraction [@prainsack2022data]. With this approach, we purposefully challenge the dominant paradigm of designing and implementing centralized platforms to support the digital transformation of healthcare in LMICs [@ogundaini2022systematic] with the aim to make digital platforms work *for* development [@hermes2020digital].

| Level | Description | Number of facilities | 
|:---:|----------------------|--------:|
| 2 | Dispensaries and private clincs, typically located in a school, industrial plant or other organization that dispenses medication and sometimes basic medical and dental treatment | 8,806 |
| 3 | Health centres, medium-sized units which cater for a population of about 80,000 people | 2,559 |
| 4 | Sub-county hospital, similar to health centres with additional facilities for more complex procedures |   971 |
| 5 | County referral hospital, regional centres which provide specialised care |    34 |
| 6 | National referral hospital | 5 |

: Number of healthcare facilities in Kenya. Source: Kenya Health Facility Census, Ministry of Health, September 2023. {#tbl-kenya-facilities}

#### SQL-on-FHIR v2 as an intermediate representation for FHIR data in tabular format

The premise of separating the user interface from the execution engine is directly related to the key objective of the SQL-on-FHIR project ([https://build.fhir.org/ig/FHIR/sql-on-fhir-v2/](https://build.fhir.org/ig/FHIR/sql-on-fhir-v2/)), namely to make large-scale analysis of FHIR data accessible to a larger audience, portable between systems and to make FHIR data work well with the best available analytic tools, regardless of the technology stack. However, to use FHIR effectively analysts require a thorough understanding of the specification as FHIR is represented as a graph of resources, with detailed semantics defined for references between resources, data types, terminology, extensions, and many other aspects of the specification.  Most analytic and machine learning use cases require the preparation of FHIR data using transformations and tabular projections from its original form. The task of authoring these transformations and projections is not trivial and there is currently no standard mechanisms to support reuse.

The solution of the SQL-on-FHIR project is to provide a specification for defining tabular, use case-specific views of FHIR data. The view definition and the execution of the view are separated, in such a way that the definition is portable across systems while the execution engine (called runners) are system-specific tools or libraries that apply view definitions to the underlying data layer, optionally making use of annotations to optimize performance.

##### Ingestion

Default workflow is extraction of data from SHR using Bulk FHIR API. Data contains metadata (incl. FHIR versions) and fully qualified semantics, for example, coding systems. Despite this, metadata extraction and metadata modeling is still required to meet the FAIR requirements. Issues that need to be solved by these services:

- To prepare for future updates of FHIR versions
- Implement late-binding principle of having increasingly more specific FHIR profiles as bulk FHIR data propagates through lakehouse

##### Storage

- File-based:
  - from ndjson to parquet
  - possibly used delta lake for time versioning
  - separation of storage from compute not only for benefits of lower TCO, but also be ready for federated learning and MPC in future

- OLAP DBMS
  - Often columnar, like Clickhouse and BigQuery
  -

##### Query & Processing

- fit in structure of OpenHIE specification
- check which workflows are related to analytics
- Hai calls this 'Maintenance'

##### Maintenance

- SQL-on-FHIR Views provide new standard to support mADX aggregate reporting !! We need to stress this, because this is an existing OpenHIE workflow
- Maintenance-related functions remain the same
- NB: orchestration falls under data provenance

##### Data consumers

- Many tools, often focus on creating information dashboards and visualizations
- Compatibility with processing & API: which query languages and interfaces are supported. Some dialect of SQL, some dialect of NoSQL, dataframe API, all of the above?

## Discussion

### Openness of health data sharing platforms

The shift in perspective from digital platforms to data platforms coincides with the paradox of open [@keller2021paradox]. Originally, openness of digital platforms focused on open source and open standards (as shown above for OpenHIE) which by has been superseded by “... conflicts about privacy, economic value extraction, the emergence of artificial intelligence, and the destabilizing effects of dominant platforms on (democratic) societies. Instead of access to information, the control of personal data has emerged in the age of platforms as the critical contention.” [@keller2021paradox]. These conflicts are particularly salient in the healthcare domain, where people are generally willing to share their health data to receive the best care (primary use, which is aligned with the concept of digital platforms), while the attitude towards secondary use of health data (conceptually aligned with a data platform) varies greatly depending on the type and context [@cascini2024health]. The shift in perspective from digital platforms supporting primary data sharing toward data platforms supporting secondary data sharing is one of the key issues surrounding the polemic of data spaces [@otto2022designing] and data solidarity [@kickbusch2021lancet;@prainsack2022data;@prainsack2023beyond;@purtova2023data].

### Conceptualizing and implementing open health data sharing platforms

In the light of these nascent paradigm shifts, the objective of this paper is the following. Given high-level design principles that we believe are supportive of realising solidarity-based, namely [TO DO: this is just first shot at formulating our design principles; needs refinement]:

- inclusive-by-design, based on the notions of datasolidarity and maximising autonomy of all future participants in the ecosystem;
- scaleable-by-design, particularly focusing on downward scaleability to support a decentralized platform topology to allows for bottom-up deployment scenarios (from local care networks --> county-level networks --> national networks) instead of top-down national roll-out;
- open-by-design, whereby a balance is found to resort to minimal standards and allow for a large diversity of partners and technologies to be used;

We specifically address the notion of openness of OHDPs in LMICs in terms of the design-related questions put forward by de Reuver at al.11:

- Object of openness: what data-related resources should data platforms make available when opening up (e.g. data, data products, datadriven insights, analytics modules)? Which user groups derive value from accessing data-related resources from data platforms (e.g. data providers, data users, intermediaries, developers)?
- Unit of analysis: what is platform-to-platform openness in the context of data platforms, given the expectation that different OHDPs will emerge at various aggregation levels? How do we distinguish meta-platforms, forking, and platform interoperability?
- Risk of openness: What are the novel (negative) implications of opening up data platforms? How can reflexivity in design help providers to resolve the negative implications of openness?
- Answers/insights to above:
  - Openness of standardized view on FHIR data and cross-language serialization of relational algebra makes it possible to fully standardize the workflow from start to finish
  - Platform-to-platform: MPC
  - Risk of openness: difficult to answer ...
- Paradox of open in disucssion: we started with hypotheses that a decentralized approach will lead to distribution of power, hence ... But is this really the case? Will open source not backfire and strengthen their position?

### Comparison with HMIS component

- Workflow requirements: Report aggregate data (link): receiver is HMIS, mADX; this is not analytics!
- Functional requirements: https://guides.ohie.org/arch-spec/openhie-component-specifications-1/openhie-health-management-information-system-hmis

Requirements are similar, but implementation differs: Datamodel is non-FHIR, focused on DataValue, which conceptually equates to FHIR Measure

### The need to a semantic layer?

- FHIR and FAIR
  - How does FHIR relate to approaches taken by the FAIR community, which tend to take more an approach of using knowledge graphs. For example, VODAN Africa [@gebreslassie2023fhir4fair;@purnamajati2022data].
  - FAIR principles vs FHIR graph: is FHIR a FAIR Data Object

- Since we use FHIR, we don't need a semantic layer because that is already provided
- We do need different semantic layer, namely with metrics. Explain different types of semantics.
  - The metrics layer same function as CQL. Discuss CQL vs generic metrics layer.

### Attribute-based access control

- TO DO: if you have generated flattened SQL tables, how are you going to manage security?
- Cerbos, attribute based on lineage or anonymized tables
- Catalogs solve this: Tabular.io, Google BigLake. What is open source option?

### Federated learning and multiparty computation

- data stations??!

## Acknowledgements

## Conflicts of interest

DK received funding from PharmAccess to conduct this work as a contractor.
SW/IntelliSOFT received funding from PharmAccess as an implementation partner in various projects.
AP/ONA received funding from PharmAccess to develop an improved version of one of the components of the open source OpenSRP 2 framework.

## Abbreviations

|    |    |
|:---|:---|
| ACID | Atomicity, Consistency, Isolation, and Durability |
| CLI | Command-line Interface |
| CR | Client Registry |
| DHI | Digital Health Intervention |
| ELK | Elasticsearch, Logstach and Kibana stack |
| ELT | Extract, Load and Transform |
| FAIR |Findable, Accessible, Interoperable and Reusable |
| FHIR |Fast Healthcare Interoperability Resources |
| FL | Federated learning |
| HIE | Health Information Exchange |
| ICT4D | ICT for Development |
| IR | Intermediate Representation |
| LMIC | Low- and middle income countries |
| MPC | Multiparty Computation |
| PET | Privacy-enhancing technologies |
| OHDS | Open health data system |
| SHR | Shared Health Record |